Local Spectral Clustering of Density Upper Level Sets

### 2019-06-24 neurips/local_spectral_density_clustering_arxiv.tex
very minor updates
halfway through conductance lower bound reorganization

\documentclass[11pt,twoside]{article}
\usepackage{fancyhdr}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue,bookmarks=false]{hyperref}
\usepackage{amsfonts,epsfig,graphicx}
\usepackage{afterpage}
\usepackage{amsmath,amssymb,amsthm} 
\usepackage{fullpage}
\usepackage[T1]{fontenc} 
\usepackage{epsf} 
\usepackage{graphics} 
\usepackage{amsfonts,amsmath}
\usepackage[sort,numbers]{natbib} 
\usepackage{psfrag,xspace}
\usepackage{color,etoolbox}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-6cm}
\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{-4cm}
\addtolength{\textheight}{-1.1\headheight}
\addtolength{\textheight}{-\headsep}
\addtolength{\textheight}{-\footskip}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\renewcommand{\floatpagefraction}{.8}%

@@

\section{Main Results}
\label{sec: consistent_cluster_estimation_with_ppr}
In this section we present our main results on the accuracy of the PPR algorithm 
for recovering density clusters. 
We begin by formally introducing a variety of geometric conditions, 
and use these to define a condition
number $\kappa(\Cset)$, which measures the difficulty PPR will have in  
estimating a density cluster $\Cset$.
With this condition in place 
our first main result (Theorem~\ref{thm: misclassification_rate}) provides a bound
on the symmetric set difference between the estimated cluster $\Cest$,
obtained by an appropriately initialized version of the PPR algorithm,
and
the target cluster $\Cset$.
Our next main result (Theorem~\ref{thm: consistent_recovery_of_density_clusters})
shows that for sufficiently well-conditioned target clusters $\Cset$, the output 
of the PPR algorithm $\Cest$
is consistent in the sense defined in Definition~\ref{def: consistent_density_cluster_estimation}.

\subsection{Preliminaries}%: Geometric Conditions and the Condition Number of a Density Cluster}
%We formalize some geometric conditions, and use these to define a condition
%number $\kappa(\Cset)$, which measures the difficulty PPR will have in  
%estimating $\Cset$. (Our theoretical guarantees for PPR will be framed in terms 
%of $\kappa(\Cset)$.)
%\paragraph{Geometric conditions on density clusters.} 
At a high level, for PPR
to be successful, the underlying density cluster must be geometrically
well-conditioned.  A basic requirement is that we need to avoid clusters which contain arbitrarily
thin bridges or spikes. As in the work of \citet{chaudhuri2010}, we consider a
thickened version of \smash{$\Cset \in \Cbb_f(\lambda)$} defined as 
$\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq \sigma}$, which 
we call the \emph{$\sigma$-expansion} of $\Cset$. Here 
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

@@

\item 
  \label{asmp: low_noise_density}
  \emph{Low noise density:} There exist $\gamma,c_0 > 0$ such that for 
  $x \in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,   
  \begin{align*}
  \inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \dist(x, \Csig)^{\gamma}. 
  \end{align*}
  Roughly, 
  this assumption ensures that the density decays sufficiently quickly as we move away from
  the target cluster $\Csig$, and is a standard assumption in the level-set estimation literature
  (see for instance~\cite{singh2009}).

  %%% AJG 5/20: Should I turn this into two conditions?
\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists $g: \Reals^d \to \Reals^d$ with the following properties: 
  \begin{enumerate}
  \item we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K} \subseteq \Rd$
  with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} =:
  \diam < \infty$
  \item $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where
  $\nabla g(x)$ is the Jacobian of $g$ evaluated at $x$ and
  \item for some $L
  \geq 1$,   
  \begin{equation*}
    \frac{1}{L}\norm{x - y} \leq \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
  \end{equation*}
  \end{enumerate}
  Succinctly, we assume that $\Csig$ is the image of a convex set with finite diameter 
  under a  measure preserving, bi-Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  We assume that neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
  \begin{equation*}
    2 \int_{\Csig} \Pbb(B(x,r)) f(x) dx \leq \int_{\Rd} \Pbb(B(x,r)) f(x) dx,
  \end{equation*}
  where $B(x,r)$ is the closed ball of radius $r$ at $x$.
\end{enumerate}
To motivate these conditions, we provide a brief high-level sketch
of our analysis (we return to this more formally in Section~\ref{sec: analysis}).
\citet{zhu2013} show for an arbitrary graph $G =
(V,E)$ and subset of vertices $S \subseteq V$, the PPR estimate \smash{$\Cest$}
of subset $S$ satisfies, for a constant $c>0$,
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\Cest \setminus S; G) + \vol(S \setminus \Cest; G) \leq c \bigl( \Phi(S,G)
\cdot \tau_{\infty}(G[S])\bigr) \vol(S;G),
\end{equation}
where $\Phi(S;G)$ is the normalized cut of $S$ (as defined in \eqref{eqn:
  normalized_cut}), and \smash{$\tau_{\infty}(G[S])$} is  the \emph{mixing
  time} of a random walk over the induced subgraph $G[S]$ (to be defined
precisely later, in \eqref{eqn: mixing_time}).  The left-hand side in
\eqref{eqn:  graph_symmetric_set_difference_1} resembles a (degree-weighted) 
form of the symmetric set difference metric in \eqref{eqn:
  misclassification_rate}, and our main goal will be to upper bound 
  this quantity. 
  Towards this goal, as we will show in Section \ref{sec:
  analysis}, the conditions \ref{asmp: bounded_density}--\ref{asmp:
  bounded_volume} allow us to upper bound the normalized cut 
\smash{$\Phi(\Csig[\Xbf]; G_{n,r})$}, and the mixing time 
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. In more detail, assumptions  
\ref{asmp: cluster_separation}, \ref{asmp: low_noise_density} yield an upper
bound on \smash{$\cut(\Csig[\Xbf]; G_{n,r})$}, and \ref{asmp: bounded_density}
yields a lower bound on \smash{$\vol(\Csig[\Xbf]; G_{n,r})$}; together with 
\ref{asmp: bounded_volume}, this gives an upper bound on the normalized cut.  On 
the other hand, \ref{asmp: bounded_density}, \ref{asmp: embedding} preclude
bottlenecks in the induced subgraph \smash{$G_{n,r}[\Csig[\Xbf]]$}, and combined
with the upper bound on diameter in \ref{asmp: embedding}, this leads to an upper bound on
the mixing time over this subgraph.

\subsubsection{Condition number} 
Motivated by \eqref{eqn:
  graph_symmetric_set_difference_1}, we will define $\kappa(\Cset)$ to be an
upper bound on the product \smash{$\Phi(\Csig[\Xbf]; G_{n,r}) \cdot
  \tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. Following, the analysis of \citet{zhu2013}
  we see that the smaller $\kappa(\Cset)$ is, the
more success PPR will have in recovering the target cluster $\Cset$. We let $\theta := (r, \sigma,
\lambda, \lambda_{\sigma}, \Lambda_{\sigma}, \gamma, \diam, L)$ contain the
geometric parameters from Assumptions~\ref{asmp:  bounded_density}--\ref{asmp:
  bounded_volume}.

\begin{definition}[Well-conditioned density clusters]
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy 
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some
  $\theta$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified
  later, we set 
  \begin{equation}
    \label{eqn: condition_number_1}
    \Phibf(\theta) 
    := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}},~  
    \taubf(\theta) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
      L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3,
  \end{equation}
  and letting \smash{$\kappa(\Cset) := \Phi_{u}(\theta) \cdot
    \tau_{u}(\theta)$}, we call $\Cset$ a 
  \emph{$\kappa$-well-conditioned density cluster}.  
\end{definition}
\noindent The condition number 
$\kappa(\Cset)$ succinctly captures the role of the various
geometric parameters. 
We note in passing that $\Phibf(\theta)$ and $\taubf(\theta)$ are exactly the upper bounds 
on \smash{$\Phi(\Csig[\Xbf]; G_{n,r})$} and
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in our analysis
later, in Section \ref{sec: analysis}.

\subsubsection{Well-initialized algorithm} As is typical in the local clustering
literature, our algorithmic results will be stated with respect to specific
ranges of each of the user-specified parameters. In particular, for a
well-conditioned density cluster $\Cset$ (with respect to some $\theta$), we
require that the tuning parameters of Algorithm~\ref{alg: ppr} are chosen as
\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in [1/10, 1/9] \cdot \frac{1}{\taubf(\theta)}, \\
v \in \Csig[\Xbf]^g, \quad
\vol_0 \in [3/4,5/4] \cdot n(n-1) \int_{\Csig} \Pbb(B(x,r)) f(x) dx, 
\end{gathered}
\end{equation}
where \smash{$\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$} will be some large
(``good'') subset of $\Csig[\Xbf]$. In particular, abbreviating $\vol_{n,r}(S)
:= \vol(S; G_{n,r})$ for $S \subseteq \Xbf$, this good subset has large volume, i.e. 
$\vol_{n,r}(\Csig[\Xbf]^g) \geq \vol_{n,r}(\Csig[\Xbf])/2$.

\begin{definition}
If the input parameters to Algorithm \ref{alg: ppr} satisfy \eqref{eqn:
  initialization} for some well-conditioned density cluster $\Cset$, we
say the algorithm is \emph{well-initialized}. 
\end{definition}
\noindent In practice it is  not feasible to set hyperparameters based on the
underlying (unknown) density $f$. Typically, one tunes PPR over a range of
hyperparameters and selects the cluster which has the smallest normalized cut. 
%it is not obvious how this scheme would affect the performance of PPR in the
%density clustering context.

\subsection{Cluster Recovery in Symmetric Set Difference}
The results of Section \ref{sec: analysis}, combined 
with \eqref{eqn: graph_symmetric_set_difference_1}, give an upper bound on the
volume of \smash{$\Cest \setminus \Csig[\Xbf]$} and \smash{$\Csig[\Xbf]
  \setminus \Cest$},  
\begin{equation}
\label{eqn: graph_symmetric_set_difference}
\vol_{n,r}(\Cest \setminus \Csig[\Xbf]) + \vol_{n,r}(\Csig[\Xbf] \setminus
\Cest) \leq c \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf]). 
\end{equation}
To translate the guarantee in~\eqref{eqn: graph_symmetric_set_difference} into meaningful bounds
on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf], \Cest)$}, we
want to preclude vertices $x \in \Xbf$ from having arbitrarily small degree, and
so we make some regularity assumptions on $\mathcal{X} := \mathrm{supp}(f)$. Let
$\nu$ denote the Lebesgue measure on $\Rd$, and $\nu_d := \nu(B)$ be the measure
of the unit ball $B = B(0,1)$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{5}
\item 
  \label{asmp: valid_region}
  \emph{Regular support:} There exists some constant $\lambda_{\min} > 0$ such 
  that $\lambda_{\min} < f(x)$ for all $x \in \mathcal{X}$. Additionally, there
  exists some $c > 0$ such that for each $x \in \partial \mathcal{X}$,
  $\nu(B(x,r) \cap \mathcal{X}) \geq c \nu_d r^d$. 
\end{enumerate}
We note that the latter condition in \ref{asmp: valid_region} will hold if the
boundary $\partial \mathcal{X}$ is sufficiently regular.  Now we present our
first main result: a bound on the symmetric set difference metric.
\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume that $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >
  0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}
The proof of Theorem \ref{thm: misclassification_rate} is deferred to Appendix~\ref{sec: proof_of_misclassification_rate}.
This result shows that the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

\subsection{Consistent Cluster Recovery}
Neither \eqref{eqn: graph_symmetric_set_difference} nor \eqref{eqn:
  misclassification_rate_ub} imply consistent density cluster estimation in the 
sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of
consistency requires a uniform bound over $\pbf$: for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\frac{p_{w}}{\Dbf_{ww}} \leq \frac{1}{40\vol_0} < \frac{1}{11\vol_0} \leq
\frac{p_u}{\Dbf_{uu}}, 
\end{equation}
so that any sweep cut $S_{\beta}$ for $\beta \vol_0 \in [1/40,1/11]$ (i.e., any
sweep cut considered by Algorithm \ref{alg: ppr}) will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that such a gap exists with probability
1 as $n \to \infty$, and hence guarantees \smash{$\Cest$} will be a consistent  
estimator. As was the case before, we wish to preclude arbitrarily low degree
vertices, this time for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >  0$
  such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{\lambda_{\sigma}^2r^d
      \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)}, 
  \end{equation}
  then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for
  $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}
\noindent Some remarks are in order: 
 \begin{enumerate}  
 \item We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in a symmetric set difference \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound in~\eqref{eqn: ppr_gap}. 
 \item While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring that the graph $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ unfortunately prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is standard for spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{enumerate}
\sbcomment{You are using $\epsilon$ here and in the future to mean different things. I have
changed this $\epsilon$ to $\varepsilon$.}

>>>>

\documentclass[11pt,twoside]{article}
\usepackage{fancyhdr}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue,bookmarks=false]{hyperref}
\usepackage{amsfonts,epsfig,graphicx}
\usepackage{afterpage}
\usepackage{amsmath,amssymb,amsthm} 
\usepackage{fullpage}
\usepackage{epsf} 
\usepackage{graphics} 
\usepackage{amsfonts,amsmath}
\usepackage[sort,numbers]{natbib} 
\usepackage{psfrag,xspace}
\usepackage{color,etoolbox}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-6cm}
\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{-4cm}
\addtolength{\textheight}{-1.1\headheight}
\addtolength{\textheight}{-\headsep}
\addtolength{\textheight}{-\footskip}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}

@@

\section{Estimation of well-conditioned density clusters}
\label{sec: consistent_cluster_estimation_with_ppr}

We formalize some geometric conditions, and use these to define a condition
number $\kappa(\Cset)$, which measures the difficulty PPR will have in  
estimating $\Cset$. (Our theoretical guarantees for PPR will be framed in terms 
of $\kappa(\Cset)$.)

\paragraph{Geometric conditions on density clusters.} At a high level, for PPR
to be successful, the underlying density cluster must be geometrically
well-conditioned.  At a minimum, we want to avoid sets that contain arbitrarily
thin bridges or spikes.  Hence, as in \citet{chaudhuri2010}, we consider a
thickened version of \smash{$\Cset \in \Cbb_f(\lambda)$} defined as 
$\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq \sigma}$, which 
we call the \emph{$\sigma$-expansion} of $\Cset$. Here 
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

@@

\item 
  \label{asmp: low_noise_density}
  \emph{Low noise density:} There exist $\gamma,c_0 > 0$ such that for 
  $x \in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,   
  $\inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \dist(x, \Csig)^{\gamma}$.

  %%% AJG 5/20: Should I turn this into two conditions?
\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists $g: \Reals^d \to \Reals^d$ with the following properties: i)
  we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K} \subseteq \Rd$
  with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} =:
  \diam < \infty$; ii) $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where
  $\nabla g(x)$ is the Jacobian of $g$ evaluated at $x$; and iii) for some $L
  \geq 1$,   
  \begin{equation*}
    \frac{1}{L}\norm{x - y} \leq \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
  \end{equation*}
  Succintly, $\Csig$ is the image of a convex set with finite diameter 
  under a  measure preserving, bi-Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  Let the neighborhood graph radius $0 < r \leq \sigma/2d$ be such that
  \begin{equation*}
    2 \int_{\Csig} \Pbb(B(x,r)) f(x) dx \leq \int_{\Rd} \Pbb(B(x,r)) f(x) dx,
  \end{equation*}
  where $B(x,r)$ is the closed ball of radius $r$ at $x$.
\end{enumerate}

To motivate these conditions, \citet{zhu2013} show for arbitrary graph $G =
(V,E)$ and subset of vertices $S \subseteq V$, the PPR estimate \smash{$\Cest$}
of subset $S$ satisfies, for a constant $c>0$,
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\Cest \setminus S; G) + \vol(S \setminus \Cest; G) \leq c \bigl( \Phi(S,G)
\cdot \tau_{\infty}(G[S])\bigr) \vol(S;G),
\end{equation}
where $\Phi(S;G)$ is the normalized cut of $S$ (as defined in \eqref{eqn:
  normalized_cut}), and \smash{$\tau_{\infty}(G[S])$} is called the \emph{mixing
  time} of a random walk over the induced subgraph $G[S]$ (to be defined
precisely later, in \eqref{eqn: mixing_time}).  The left-hand side in
\eqref{eqn:  graph_symmetric_set_difference_1} resembles a (degree-weighted) 
form of the symmetric set difference metric in \eqref{eqn:
  misclassification_rate}.  As we will show in Section \ref{sec:
  analysis}, the conditions \ref{asmp: bounded_density}--\ref{asmp:
  bounded_volume} allow us to upper bound the normalized cut 
\smash{$\Phi(\Csig[\Xbf]; G_{n,r})$}, and the mixing time 
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}.\footnote{Informally, assumptions  
\ref{asmp: cluster_separation}, \ref{asmp: low_noise_density} yield an upper
bound on \smash{$\cut(\Csig[\Xbf]; G_{n,r})$}, and \ref{asmp: bounded_density}
yields a lower bound on \smash{$\vol(\Csig[\Xbf]; G_{n,r})$}; together with 
\ref{asmp: bounded_volume}, this gives an upper bound on the normalized cut.  On 
the other hand, \ref{asmp: bounded_density}, \ref{asmp: embedding} preclude
bottlenecks in the induced subgraph \smash{$G_{n,r}[\Csig[\Xbf]]$}, and combined
with the upper bound on diameter in \ref{asmp: embedding}, this leads to an upper bound on
the mixing time over this subgraph.}

\paragraph{Condition number.} Motivated by \eqref{eqn:
  graph_symmetric_set_difference_1}, we will define $\kappa(\Cset)$ to be an
upper bound on the product \smash{$\Phi(\Csig[\Xbf]; G_{n,r}) \cdot
  \tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. The smaller $\kappa(\Cset)$ is, the
more success PPR will have in recovering $\Cset$. Let $\theta := (r, \sigma,
\lambda, \lambda_{\sigma}, \Lambda_{\sigma}, \gamma, \diam, L)$ contain the
geometric parameters from \ref{asmp:  bounded_density}--\ref{asmp:
  bounded_volume}.

\begin{definition}[Well-conditioned density clusters]
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy 
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some
  $\theta$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified
  later, we set 
  \begin{equation}
    \label{eqn: condition_number_1}
    \Phibf(\theta) 
    := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}},~  
    \taubf(\theta) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
      L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3,
  \end{equation}
  and letting \smash{$\kappa(\Cset) := \Phi_{u}(\theta) \cdot
    \tau_{u}(\theta)$}, we call $\Cset$ a 
  \emph{$\kappa$-well-conditioned density cluster}.  
\end{definition}

We note that $\Phibf(\theta)$ and $\taubf(\theta)$ are exactly the upper bounds 
on \smash{$\Phi(\Csig[\Xbf]; G_{n,r})$} and
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in our analysis
later, in Section \ref{sec: analysis}.

\paragraph{Well-initialized algorithm.} As is typical in the local clustering
literature, our algorithmic results will be stated with respect to specific
ranges of each of the user-specified parameters. In particular, for a
well-conditioned density cluster $\Cset$ (with respect to some $\theta$), we
require 
\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in [1/10, 1/9] \cdot \frac{1}{\taubf(\theta)}, \\
v \in \Csig[\Xbf]^g, \quad
\vol_0 \in [3/4,5/4] \cdot n(n-1) \int_{\Csig} \Pbb(B(x,r)) f(x) dx, 
\end{gathered}
\end{equation}
where \smash{$\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$} will be some large
(``good'') subset of $\Csig[\Xbf]$. In particular, abbreviating $\vol_{n,r}(S)
:= \vol(S; G_{n,r})$ for $S \subseteq \Xbf$, we will have
\smash{$\vol_{n,r}(\Csig[\Xbf]^g) \geq \vol_{n,r}(\Csig[\Xbf])/2$}.

\begin{definition}
If the input parameters to Algorithm \ref{alg: ppr} satisfy \eqref{eqn:
  initialization} for some well-conditioned density cluster $\Cset$, we
say the algorithm is \emph{well-initialized}. 
\end{definition}

In practice it is clearly not feasible to set hyperparameters based on the
underlying (unknown) density $f$. Typically, one tunes PPR over a range of
hyperparameters and optimizes for some criterion such as minimum normalized cut;
it is not obvious how this scheme would affect the performance of PPR in the
density clustering context.

\paragraph{Main theorems.} The results of Section \ref{sec: analysis}, combined 
with \eqref{eqn: graph_symmetric_set_difference_1}, give an upper bound on the
volume of \smash{$\Cest \setminus \Csig[\Xbf]$} and \smash{$\Csig[\Xbf]
  \setminus \Cest$},  
\begin{equation}
\label{eqn: graph_symmetric_set_difference}
\vol_{n,r}(\Cest \setminus \Csig[\Xbf]) + \vol_{n,r}(\Csig[\Xbf] \setminus
\Cest) \leq c \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf]). 
\end{equation}
To translate \eqref{eqn: graph_symmetric_set_difference} into meaningful bounds
on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf], \Cest)$}, we
want to preclude vertices $x \in \Xbf$ from having arbitrarily small degree, and
so we make some regularity assumptions on $\mathcal{X} := \mathrm{supp}(f)$. Let
$\nu$ denote the Lebesgue measure on $\Rd$, and $\nu_d := \nu(B)$ be the measure
of the unit ball $B = B(0,1)$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{5}
\item 
  \label{asmp: valid_region}
  \emph{Regular support:} There exists some constant $\lambda_{\min} > 0$ such 
  that $\lambda_{\min} < f(x)$ for all $x \in \mathcal{X}$. Additionally, there
  exists some $c > 0$ such that for each $x \in \partial \mathcal{X}$,
  $\nu(B(x,r) \cap \mathcal{X}) \geq c \nu_d r^d$. 
\end{enumerate}
Note that the latter condition in \ref{asmp: valid_region} will hold if the
boundary $\partial \mathcal{X}$ is sufficiently regular.  Now we present our
main bound on the symmetric set difference metric.

\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume that $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >
  0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}

The proof of Theorem \ref{thm: misclassification_rate}, along with all other
proofs in this paper, is deferred to the appendix. Note that this 
result says the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

Neither \eqref{eqn: graph_symmetric_set_difference} nor \eqref{eqn:
  misclassification_rate_ub} imply consistent density cluster estimation in the 
sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of
consistency requires a uniform bound over $\pbf$: for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\frac{p_{w}}{\Dbf_{ww}} \leq \frac{1}{40\vol_0} < \frac{1}{11\vol_0} \leq
\frac{p_u}{\Dbf_{uu}}, 
\end{equation}
so that any sweep cut $S_{\beta}$ for $\beta \vol_0 \in [1/40,1/11]$ (i.e., any
sweep cut considered by Algorithm \ref{alg: ppr}) will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that such a gap exists with probability
1 as $n \to \infty$, and hence guarantees \smash{$\Cest$} will be a consistent  
estimator. As was the case before, we wish to preclude arbitrarily low degree
vertices, this time for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >  0$
  such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{\lambda_{\sigma}^2r^d
      \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)}, 
  \end{equation}
  then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for
  $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}

\begin{remark}
  We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in a symmetric set difference \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound of \eqref{eqn: ppr_gap}. 
\end{remark}

\begin{remark}
  While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ unfortunately prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is standard for spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{remark}
